% vim:ft=tex
% rubber: module xelatex

\section{Conclusions}
\label{sec:conclusions}
We have implemented image rectification based on the our earlier work
in camera calibration, using the intrinsic and extrinsic parameters from the
calibration to compute the rectification. We have also implemented a
dynamic programming based stereo matching algorithm for creating depth maps from
the rectified images.

Together, these formed the foundation for our implementation of PCA-based
holistic face recognition. PCA uses up to six channels of data from the images
(RGB, hue, saturation and depth data), and works with considerable accuracy. We
were pleasantly suprised by the success of PCA on even very small face images,
such as $32\times49$ pixels. However, our testing was limited by the rather
homogeneous face database we used (all young males in the same lighting and
image positions).

We have tested our implementation of the different parts and provide accuracy
test results for the different parts. As part of our testing efforts, we have
also improved on the testing of our calibration implementation, and provided
additional results at the beginning of this report.

Overall, our test results show that we achieve reasonable results in each part
of our implementation, but there is room for improvement. We need to find a way
to create depth maps \emph{in reasonable time} on images that have been
rectified but \emph{not} scaled and cropped by preprocessing, as this appears to
reduce their quality. Our stereo matching and rectification could in general be
improved in quality, and the efficiency and speed of our face recognition
algorithms could be increased.
