% vim:ft=tex
% rubber: module xelatex
\subsection{Stereo matching}
\label{sec:stereo-prior}

Stereo matching, the process of reconstructing a 3D scene from two (or possibly more) stereo
images, is an important part of computer vision. Various approaches to the
problem have been proposed.

Stereo matching algorithms typically fall into two categories: those doing local
matching (i.e. per-pixel or in pixel neighbourhoods) and those using a global
matching model (i.e. analysing the whole image). The global method requires
keeping track of a lot of information about the whole image, which can be quite
demanding on computational resources (mainly memory), while the local method
requires selecting an appropriate matching window, which may vary between
images. An evaluation of a large number of different algorithms is given by
\citet{stereocorrespondence}.

One example of the local matching approach is the correlation-based matching
algorithm with a fast GPU implementation given by \citet{weber09}. Global
matching approaches include the use of graph cuts
\cite{kolmogorov01:_comput}, and belief propagation using Markov fields
\cite{felzenszwalb06:_effic_belief_propag_early_vision}.

Improving the \emph{speed} of stereo matching is the focus of
\citet{geiger11:_effic_large_scale_stereo_match}, who achieve close to real-time
matching speeds with high accuracy using a set of support points that can be
accurately identified between the two images. Another real-time approach is taken
by \citet{realtimestereo}, who apply dynamic programming techniques for rapid stereo matching.

Real-time stereo vision is important in domains such as robotics, where information needs to be integrated over time as well as space. For instance, moving objects need to be tracked. The spatial windows of block-correlation stereo matching can be extended to include a time dimension: a `spatio-temporal' window aggregating information at each pixel. \citet{tombari_konolige} demonstrate an efficient innovation for spatial regularisation, \emph{local smoothing}.
